---
title: 'Team CarJan: Predict tree cover'
author: "Jan Droesen and Careli Caballero"
date: "January 14, 2016"
output: html_document
---
In this report a model is developed for predicting the percentage of tree cover at the Gewata dataset. 
```{r setup, include=FALSE}
## libraries
library(raster)
library(rgdal)
library(graphics)

## load data
load("data/GewataB2.rda")
load("data/GewataB3.rda")
load("data/GewataB4.rda")
load("data/GewataB1.rda")
load("data/GewataB5.rda")
load("data/GewataB7.rda")
load("data/vcfGewata.rda")
```

## Processing data and plot bands against tree cover

```{r bands and VCF}

#Erase tree cover higher than 100%
vcfGewata[vcfGewata > 100] <- NA

#make bricks and assign names to each brick layer
bandsgewata <- brick(GewataB1,GewataB2, GewataB3, GewataB4, GewataB5, GewataB7)
alllayers <- addLayer(bandsgewata, vcfGewata)
names(bandsgewata) <- c("BlueBand", "GreenBand", "RedBand", "NIR", "SWIR", "MIR")
names(alllayers) <- c("BlueBand", "GreenBand", "RedBand", "NIR", "SWIR", "MIR", "VCF")

#plot all bands with VCF
par(mfrow = c(3,2))
plot(alllayers[[1:6]], alllayers[[7]], xlab = names(alllayers), ylab = ("Tree cover (%)"))
```

Each plot represents the relation between LANDSAT7 ETM+ bands and the Vegetation Continuous Field product. High correlations for the red and SWIR can be derived.

```{r Linear model (LM)}
#process data for linear regression model
dfgewata <- as.data.frame(getValues(alllayers))
#erase NAś
dfgewata <- na.omit(dfgewata)

#multiple regression model with all bands
modelMultipleAll <- lm(VCF ~ BlueBand + GreenBand + RedBand + NIR + SWIR + MIR, data = dfgewata)
#check for significant bands
summary(modelMultipleAll)

#erase non-significant band(s) and run linear regression model
modelMultiple <- lm(VCF ~ BlueBand + GreenBand + RedBand + NIR + SWIR, data = dfgewata)
```
##
From the multiple regression model of all bands the signifance is investigated. Non-significant bands (i.e. only the Middle Infrared (MIR) band) were erased. The optimal multiple regression model is now calculated.

```{r comparison LM and VCF}


#predicted values
predTC <- predict(bandsgewata, model=modelMultiple, na.rm=TRUE)

#comparing predicted and VCF tree cover

#get the difference between VCF and predicted
raserDif <- vcfGewata - predTC

#calculate RMSE separately for each of the classes
calcRMSE <- function(x,y){
	sqrt(mean((x-y)^2))
}
polyError <- calcRMSE(vcfGewata,predTC)

#visualize difference beteen VCF and predicted
colfunc <- colorRampPalette(c("white", "red"))
par(mfrow = c(2,2))
plot(vcfGewata, main="Tree cover (VCF)", zlim=c(0,100) )
plot(predTC, main="Tree cover predicted (LM)", zlim=c(0,100)) #add 0-100
plot(raserDif, main="Difference in tree cover", col=colfunc(255),  zlim=c(0,50))
plot(polyError, main="RMSE of VCF and LM",col=colfunc(255),  zlim=c(0,50))
```

In the plots the tree cover from the VCF product and the linear model is shown. Remarkable is that in the areas with high tree cover the errors seems to be higher. To check this the errors of the LM are calculated for three land cover types.

```{r error LM for land cover}
#get errors for the LM with the land cover types
load("data/trainingPoly.rda")
trainingPoly@data$Code <- as.numeric(trainingPoly@data$Class)
classes <- rasterize(trainingPoly, polyError, field='Code')

#calculate error per land cover class
errorTraining <- zonal(polyError, classes, fun='mean', digits=0)
dferrorTraining <- as.data.frame(errorTraining)

#which class corresponds to highest error
mergedPolyError <- merge(x = trainingPoly, y = dferrorTraining, by.x = "Code", by.y = "zone")

#visualize errors per land cover type
x <- mergedPolyError$Class
y <- mergedPolyError$mean
par(mfrow=c(1,1))
plot(x,y, ylim= c(0,10), "s", ylab="RMSE")
```

After getting the error per each land cover class (cropland, forest and wetland), the most accurate class was forest with the lowest error, which was as expected, because of the VCF product.